---
title: '50 Days Of Food'
author: "Will Bidstrup"
date: "06/09/2021"
output: html_document
---

This is the companion notebook to my experiment visualizing 50 Days Of Food.

# Setup

I am a fan of the `tidyverse` so I will load it in.  

```{r}

install.packages("GGally")
install.packages("svglite")
```


```{r}
# Packages
library(tidyverse)
library(GGally)
library(svglite)

```

# Data ingestion

MyFitnessPal export generates three files, for Exercise, Measurements, and Nutrition. I export these from the app to my email and then put them in the 'data' folder of my project.  

```{r}
exercise <- read_csv("data/Exercise-Summary-2021-07-19-to-2021-09-06.csv")
measurement <- read_csv("data/Measurement-Summary-2021-07-19-to-2021-09-06.csv")
nutrition <- read_csv("data/Nutrition-Summary-2021-07-19-to-2021-09-06.csv")
```


# Data exploration

## Nutrition file

The nuttiotion file is the most interesting one so I will start here and only move to the others if I have an idea that requires them (for example using steps).  

I really like the `glimpse` function from `tidyverse`

```{r}
# Inspect nutrition file

glimpse(nutrition)
```

The nutrition file has 21 columns and is already in pretty good shape. The date is already ISO formatted, there are lots of numeric data. This is ready to go.  


# Data visualisation

For me, data visualization almost always involves some level of data transformation as well. Only onece I have found something worth keeping will I move ETL to a seperate process.


## Energy consumption over time

This will be a basic line graph of energy consumption over time.

```{r}
# Visualise
ggplot(data = nutrition, aes(x = Date, y = Kilojoules)) +
  geom_line()
```

You can see from above that something doesn't quite look right. It's because the Kilojoules aren't aggregated by day, but by meal (Breakfast, Snacks, Lunch, Dinner).

Some pre-processing will help, I always make it explicit when I am doing exploratory work as below:  

```{r}

# Pre-process
kj_by_day <- nutrition %>%
  group_by(Date) %>%
  summarise(Daily_Kilojoules = sum(Kilojoules, na.rm = FALSE))

# Visualise
ggplot(data = kj_by_day, aes(x = Date, y = Daily_Kilojoules)) +
  geom_line() +
  theme_bw()

```

## Salt 

R is great for scatter plots ans visualizing lots of points. I wonder if looking at salt will be interesting. 


```{r}
glimpse(nutrition)
```


```{r}
salt_analysis <- nutrition %>%
  select(Kilojoules, `Sodium (mg)`)

# Visualise 

ggplot(data = nutrition, aes(x = Kilojoules, y = `Sodium (mg)`)) +
  geom_point()
```
Now this is interesting, looks like I had 5000mg of sodium in one meal? 


How about cumulative salt?

```{r}
cum_salt <- nutrition %>%
  select(Date, `Sodium (mg)`) %>%
  mutate(cum_salt = cumsum(`Sodium (mg)`))
```

This could be a cool visual...I have consumed about 80g of salt.  

```{r}
ggplot(data = cum_salt, aes(x = Date, y = cum_salt)) +
  geom_line() +
  theme_bw()
```



## Text analysis, most common foods

It turns out there is no text information shared by default, it needs to be added manually in notes. This seems like a big miss.  


## Correlation of nutrients

```{r}
ggcorr(nutrition)
```



```{r, fig.height = 10, fig.width = 10}
nutrition_numeric <- nutrition %>%
  select(- c(Date, Meal, Time, Note))

ggpairs(nutrition_numeric)
```

## Hero Visual 1: Summary of Meals


I want a visual which shows a lot of things:  

- Time of day  
- Meal  


```{r}
ggplot(data = nutrition %>%
         mutate(Meal = fct_relevel(Meal,"Snacks", "Dinner", "Lunch", "Breakfast")), aes(x = Time, y = Meal)) +
  geom_boxplot() +
  geom_point(position = "jitter", alpha = 0.2) +
  theme_bw()
  
```

Some pre-processing will help here

I want to go top to bottom breakfast, lunch, dinner, snacks
```{r}
# Re-order factors

ggplot(data = nutrition %>%
         mutate(Meal = fct_relevel(Meal,"Snacks", "Dinner", "Lunch", "Breakfast")), aes(x = Time, y = Meal)) +
  geom_boxplot() +
  geom_point(position = "jitter", alpha = 0.5) +
  theme_minimal() +
  labs(x = "",
       y = "")



```

```{r}
quantile_kj <- quantile(nutrition$Kilojoules,c(0,1/3,2/3,1))
```

```{r}


nutrition <- nutrition %>%
  mutate(KJ_quantile = cut(Kilojoules, breaks = quantile_kj, labels=c("small","medium","big")),
         KJ_score = ifelse(KJ_quantile == "small", 1,
                           ifelse(KJ_quantile == "medium", 2, 3)))

nutrition$KJ_score <- as.integer(nutrition$KJ_score)
```



```{r}
ggplot(data = nutrition %>%
         mutate(Meal = fct_relevel(Meal,"Snacks", "Dinner", "Lunch", "Breakfast")), aes(x = Time, y = Meal, col = KJ_score, size = Kilojoules)) +
  geom_boxplot() +
  geom_point(position = "jitter", alpha = 0.4) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(x = "",
       y = "") +
scale_color_gradient(low="#6c5d3f", high="#ffc700")

ggsave("fig1.svg")

```

```{r}
ggplot(data = cum_salt, aes(x = Date, y = cum_salt)) +
  geom_line() +
  theme_minimal() +
  labs(x = "",
       y = "")

ggsave("fig2.png")
```



```{r}

# Pre-process
kj_by_day <- nutrition %>%
  group_by(Date) %>%
  summarise(Daily_Kilojoules = sum(Kilojoules, na.rm = FALSE))

# Visualise
ggplot(data = kj_by_day, aes(x = Date, y = Daily_Kilojoules)) +
  geom_point() +
  geom_smooth(col = "#ffc700", method = "lm") +
  theme_minimal() +
  ylim(0, 10000) +
  labs(x = "", y = "")

ggsave("fig3.png")

```
